### 2013-03-23 eusflat2013/introduction.tex
Almost final changes.

Information systems (IS) have always tried to model parts of reality. To achieve this modelling, such systems contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}. As time is an essential aspect of many real-world objects or concepts, information systems often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}. Such temporal notions usually take the form of either instants~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Jensen1998}.

@@

Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data. Information or knowledge is usually retrieved from an IS by querying it and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).

Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too. Usually, querying an IS containing data representing temporal notions is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{Billiet2012},~\cite{JoseEnriquePons2012},~\cite{Pons2012a}. Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships~\cite{Medina1994},~\cite{Schockaert2008}. Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants). The resulting relationships are shown in figure \ref{fig:allen-relationships}. These Allen relationships are used in the presented work.

@@

The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper. Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some preliminary concepts and techniques specific to the framework under consideration, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done. In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships are compared. Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.

>>>>

Information systems (IS) have always tried to model parts of reality. To achieve this modelling, IS contain data representing properties of real-world objects or concepts~\cite{JoseEnriquePons2012},~\cite{Billiet2012}. As time is an essential aspect of many real-world objects or concepts, IS often contain data representing temporal notions which describe such temporal properties~\cite{Bolour1982},~\cite{VanderCruyssen1997}. Such temporal notions usually take the form of either instants~\cite{Jensen1998}, which can informally be seen as infinitely short `moments' or `points' in time, or time intervals~\cite{Jensen1998}.

@@

Generally, one of the most important purposes of an IS is to allow the retrieval of information or knowledge deduced from its data. Such retrieval is usually done by querying the IS and examining or analyzing the query results or by visualizing the contents of the IS, querying this visualization and examining or analyzing the resulting visualization(s).

Of course, when temporal information is represented in an IS, querying this IS may have a temporal aspect too. Usually, querying such an IS is conceptually done by specifying one or more time indications and requesting information that is in a specific relationship with these indications, where the semantics of these relationships are specifically temporal~\cite{Billiet2012},~\cite{JoseEnriquePons2012},~\cite{Pons2012a}. Thus, some existing proposals have considered groups of basic relationships between time indications used to construct and express specific temporal relationships~\cite{Medina1994},~\cite{Schockaert2008}. Notably, Allen~\cite{Allen1983} presented a reasoning framework containing all semantically usefull basic temporal relationships between time intervals (and as a special case instants). The resulting relationships are shown in figure \ref{fig:allen-relationships}. These Allen relationships are used in the presented work.

@@

The structure of this paper is as follows: section \ref{sec:general-preliminaries} presents some general preliminaries and naming and notation conventions used in this paper. Sections \ref{sec:ikc} and \ref{sec:tm} introduce the IKC and TM frameworks respectively: both sections first introduce some specific preliminary concepts and techniques, then explain how the representation of time intervals by the framework is done and finally show how the evaluation of Allen relationships between an interval subject to uncertainty and one without uncertainty is done. In section \ref{sec:proposal}, both frameworks are compared: first their approaches to representing time intervals are compared, next their approaches to evaluating Allen relationships. Finally, section \ref{sec:conclusions} presents the principal conclusions of the presented work and some possible future work.

### 2013-03-23 eusflat2013/proposal.tex
Almost final changes.

\subsection{\label{subsec:sim-dif}Similarities and differences}

In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI). Then, we will compare two ill-known time points and two uncertain time indications.

>>>>

In this section, first the approaches of the IKC and TM frameworks towards interval representation are compared. Next, their approaches towards the evaluation of Allen relationships between a crisp interval and an interval subject to uncertainty are compared.

\subsection{\label{subsec:comp-interval}Comparison of Approaches to Interval Representation}
The IKC framework uses IKTI to represent time intervals subject to uncertainty, whereas the TM framework uses UTI for this. In both approaches, to consider uncertainty about the exact CTI which is intended, uncertainty about the exact starting and ending instants of the intended interval is considered and confidence in the context of this uncertainty is expressed using two possibility distributions. In an IKTI, these possibility distributions each define one IKV. One of these then defines the IKTI's starting instant and the other defines its ending instant. In an UTI, one of these possibility distributions is directly meant to define the UTI's starting instant and the other is directly meant to define the UTI's ending instant. It is obvious that the concepts of IKTI and UTI are exactly the same, except for the explicit usage of the concept of IKV in IKTI to define and describe uncertainty about starting and ending instants. This equality implies that both approaches have the same basic restriction: they cannot represent every kind of interval subject to uncertainty imaginable. For example, imagine an ordered set of instants coinciding with $\mathbb{N}_0$ and imagine an interval subject to uncertainty in this set, where the intended interval is either $\left[2, 4\right]$ (possibility $1$) or $\left[3, 7\right]$ (possibility $1$). This interval can not be modelled by a combination of possibility distributions defining the starting and ending instants without making $\left[2, 7\right]$ and $\left[3, 4\right]$ also possible as intended intervals to some extent. This issue was first suggested about IKTI in~\cite{Billiet2012}.

Now, a major difference between the two frameworks, concerning their approaches towards the representation of time intervals in general, is that the TM framework includes visualization in its approach.

A first consequence of this is that the TM framework allows the visualization of multiple CTI in the same image plane. In theory, it should also allow the visualization of multiple UTI in the same image plane. However, if the interval space area's visualizing these UTI overlap, it is not yet researched which greyscale (or other) color and intensity each interval point in this overlapping area should have, as the appearance of such interval point should both reflect the possibility of it being the interval intended by the one UTI and the possibility of it being the interval intended by the other UTI. The advantage of the ability to visualize multiple time intervals in the same image plane is that a human observer could easily assess certain characteristics of a distribution of time intervals from an image containing their visualization. The IKC framework, on the other hand, doesn't incorporate any form of visualization and doesn't allow such human assessments.

A second consequence of this is that the TM framework requires that an UTI can be visualized, using a visualization method which actually visualizes the possibility of a CTI of being the interval intended by this UTI for every CTI that has a non-zero possibility of this. Thus, a method is required, to calculate the possibility of a given CTI of being the interval intended by an UTI based on the possibility distributions defining this UTI. This method is found by demanding that the given CTI's starting instant is the intended interval's starting instant \emph{and} that the given CTI's ending instant is the intended interval's ending instant and by determining the possibility of the conjunction of both demands using the standard possibility theory conjunction operator `minimum'. Although not necessary for the correct and consistent functioning of the TM framework, it appears to be the intention of the TM framework that possibility about the exact starting or ending instant of the interval intended by the UTI could be derived from the possibility distribution defining the possibility that a given CTI is the interval intended by the UTI. For this derivation to be consistent, the possibility distributions defining the UTI must be convex. Given an ordered set $T$ and an UTI $J = (\pi_{J_s}, \pi_{J_e})$ in $T$ with possibility $\pi_J(I)$ that a given CTI $I = \left[s_i, e_i\right]$ is the exact time interval intended by $J$, the derivations can be calculated as follows:

\begin{align}
\pi_{J_s}(s_i) = \max_{K = \left[s_i, k\right], k \in T, k > s_i}(\pi_J(K)) \nonumber \\
\pi_{J_e}(e_i) = \max_{K = \left[k, e_i\right], k \in T, k < e_i}(\pi_J(K)) \nonumber
\end{align}

With respect to this convexity demand, the IKC framework is similar: the possibility distributions defining the starting and ending IKV of an IKTI are also demanded to be convex by the IKC framework, although this appears not to be necessary for the framework to function correctly and consistently.

\subsection{\label{subsec:comp-eval}Comparison of Approaches to Allen Relationship Evaluation}
As mentioned before, a major difference between the two frameworks, concerning their approaches towards the representation of time intervals in general, is that the TM framework includes visualization in its approach. As a result, it also includes visualization in its approach towards the evaluation of Allen relationships between a CTI and an UTI.

A first consequence of this is that, given a CTI, an UTI, its URZ and their visualizations in the same image, the TM framework allows a visual, human assessment of the degree of possibility of this CTI being in an Allen relationship with this UTI, for every Allen relationship with a non-zero such degree, based on this image. Moreover, those with possibility degree equal to zero can be easily found before any calculation is done: they are the relationships not contained in the set corresponding to the URZ containing the CTI's interval point. On the other hand, to examine the respective possibilities of a given CTI of being in several different Allen relationships with a given IKTI using the IKC framework, a new collection of specific IKC and a specific aggregation of these should be constructed for every Allen relationship under consideration, allowing to calculate its exact possibility and necessity. In contrast to the TM framework, using the IKC framework a human assessment before any calculation is not possible and it is not known before any calculation which Allen relationships will result in a possibility degree equal to zero.

A second consequence is that, given an UTI and its URZ, multiple CTI can be visualized in the same image as the UTI and its URZ. Thus, the same image could provide a visual, human assessment of the possibilities with which multiple CTI are in Allen relationships with the UTI, before any calculation is done. Again, the IKC framework would need a different collection of specific IKC and a specific aggregation of these for every Allen relationship under consideration, but calculating the possibilities for several CTI to be in a given Allen relationship with a given IKTI would not provide much extra work.

A third consequence is that, in the TM framework, given an UTI, its URZ and a visualization of these in the same image, a given Allen relationship could correspond with several URZ. Thus, given a distribution of CTI, it is not easy to visually assess which CTI have a non-zero possibility of being in the given Allen relationship with the given UTI. On the other hand, assessing this using the IKC framework is impossible without any calculation, but the necessary calculations are pretty straightforward.

Although this has not been rigorously researched yet, it is the conviction of the authors that a great strength of the IKC framework lies in its modular approach towards the evaluation of temporal relationships, resulting in a flexibility and an easy handling of complex temporal relationships, while the visualization of these using the TM framework could become complex and heavy. This would give the IKC framework a major advantage over the TM framework, when used in reasoning systems like e.g. decision support systems. Moreover, in some cases the visualization step used in the TM framework could be a redundant step.

%Both above statements say something about `visual assessment'. However, to actually calculate the plausibilities in question, the IKC framework requires a rather complex numerical (?) calculation based on the set of IKC and its combination corresponding to the Allen relationship and the CTI in question. In this, the TM framework is similar, as it also requires a rather complex geometrical construction and a calculation based on this, for every Allen relationship and CTI in question, IF the crisp interval point is in an URZ corresponding with a non-singleton set of Allen relationships. Else, the plausibility for the Allen relationship in the singleton is 1, all others are 0.

%Although it has not been rigorously researched yet, it is the conviction of the authors that both frameworks can also be used to evaluate combinations of Allen relationships between crisp intervals and an UTI/IKTI. For example: `what is the possibility that $J$ is before $K$ but after $I$?'.

\subsection{\label{subsec:example}An Example}
In this section, some of the comparison results presented above will be illustrated using a simple example.

\begin{example}
Consider an ordered set of instants $T$ coinciding with $\mathbb{R}$, a CTI $I = \left[1, 3\right]$ and a time interval $J$ of which the starting instant is defined by a triangular possibility distribution on $T$ with core $\{5\}$ and support $\left[2, 7\right]$ and of which the ending instant is defined by a triangular possibility distribution on $T$ with core $\{12\}$ and support $\left[9, 15\right]$. The visualization of this example situation using the TM framework is shown in figure \ref{fig:ex}. The interval point for $I$ lies in the URZ `PM'. Thus, possibility is higher than zero for $I$ to be in a `meets', `before', or `overlaps' relationship with $J$. As the darkest point in the `overlaps' area part is very light, the darkest point in the `meets' line segment is reasonably dark and the darkest point in the `before' area part is perfectly black, it can be estimated that $I$ overlaps $J$ with a low possibility, $I$ meets $J$ with a high but non-zero possibility and $I$ is before $J$ with possibility $1$.

TODO: IKC

\end{example}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\columnwidth]{graphs/example_image.eps}
	\caption{The visualization of the example using the TM framework.}
	\label{fig:ex}
\end{figure}

%We will illustrate the calculation of one Allen Relation with an example.
%\begin{example}
%Consider a CTI $J = [1, 3]$ the two ill-known time points $S, E$ that define a time interval. 
%We want to obtain the possibility that $J$ is before the time interval.
%The points are defined by two triangular membership functions $\mu_S$, $\mu_E$.

%\begin{align}
%\label{eq:ex-point}
%\pi_S &= \mu_S = \left[2,5,7 \right] \\
%\pi_E &= \mu_E = \left[9,12,15 \right]
%\end{align}

%Then, it is possible to define both an IKTI $I$ Eq. \eqref{eq:ex-uti-ikti} and an UTI $\pi_I$ Eq. \eqref{eq:ex-uti-uti}. The intervals are illustrated in Figure \ref{fig:example}.

%\begin{figure}[h]
%   \centering
%   \includegraphics[scale=1.5]{graphs/example.eps}
%   \caption{In the top, the crisp time interval $J$ and the IKTI given by $I = [S, E]$. The ill-known constraint for the before relation is illustrated in grey color. In the lower graphic, the interval point $J$ and the UTI $\pi_I$. }
%   \label{fig:example}
% \end{figure}

%\begin{align}
%\label{eq:ex-uti-ikti}
%I = \left[S, E \right]\\
%\label{eq:ex-uti-uti}
%\mu_I = \left(\mu_S, \mu_E \right)
%\end{align}

%First we will obtain the possibility between $J$ and the UTI $\pi_I$. The relation possibly meets has three candidates: Before, Meets and Overlaps.

%\begin{align}
%  \mu_{\left(J R \pi_I \right)}(B) &= \sup_{a < e} \pi_S(a)\\
%\mu_{\left(J R \pi_I \right)}(M) &= \pi_S(e)\\
%\mu_{\left(J R \pi_I \right)}(O) &= \sup_{a > e} \pi_S(a)
%\end{align}

%\begin{align}
%\mu_{\left(J R \pi_I \right)}(B) &= 0.67\\
%\mu_{\left(J R \pi_I \right)}(M) &= 0.33\\
%\mu_{\left(J R \pi_I \right)}(O) &= 0
%\end{align}

%In order to compute the possibility  for the before relation between the crisp interval $J$ and the IKTI $I$, a ill-known constraint is needed.

%\begin{equation}
% C_1\stackrel{\triangle}{=} \left(<,X\right)
%\end{equation}

%Then the possibility is computed as follows:

%\begin{align}
%\Pos \left(C_1 (J) \right) &= \sup_{a < e} \pi_S(a)\\
%\Pos \left(C_1 (J) \right) &= 0.67 
%\end{align}

%\end{example}

% 
% In this section we will analyze first the calculation of all the possible Allen's relations between a crisp time point and an ill-known time point and between a crisp time point and an uncertain time indication (UTI). Then, we will compare two ill-known time points and two uncertain time indications.
